---
title: "writing_functions"
author: "Fang Wang"
date: "2024-10-23"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r}
library (tidyverse)
library(rvest)
setwd("/Users/fangwang/Downloads/P8105 Data Science I/Lecture 16/writing_functions")

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp =.6,
  out.width = "90%"
)

theme_set (theme_bw() + theme (
    legend.position =  "bottom"
  ))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_color_discrete = scale_color_viridis_d()
scale_fill_discrete = scale_fill_viridis_d()
```

# Do sth simple: z-score computation (input: 1 vector and output: 1 vector)
```{r}
x_vec = rnorm(25, mean = 10, sd = 3.5)

(x_vec - mean(x_vec)) / sd(x_vec) # z-score computation
```

# Write a function to compute z-scores
```{r}
z_scores = function(x) {
  z = (x - mean(x)) / sd(x)
  return(z)
}

z_scores(x_vec)
```

# try my function on some other things
```{r, error = TRUE}
z_scores(3) # can't compute sd or mean from just one number!
z_scores("my name is jeff") # it is not numeric.
z_scores(iris)

z_scores(sample(c(TRUE, FALSE), 25, replace = TRUE))
```


```{r}
z_scores = function(x) {
  
  if (!is.numeric(x)) {
    stop("Argument x should be numeric") # for z_scores("my name is jeff")  
  } else if (length(x) == 1) {           # for z_scores(3)
    stop("Z scores cannot be computed for length 1 vectors")
  }
  z = mean(x) / sd(x)
  z
}


z_scores = function(x) {
  
  if (!is.numeric(x)) {
    stop("Argument x should be numeric") # for z_scores("my name is jeff")  
  } else if (length(x) < 5) {           # for z_scores(3)
    stop("you need at least five numbers to compute the z scores")
  }
  z = mean(x) / sd(x)
  z
}
```

# multiple outputs
```{r}
mean_and_sd = function(x) {
  
  if (!is.numeric(x)) {
    stop("Argument x should be numeric")
  } else if (length(x) == 1) {
    stop("Cannot be computed for length 1 vectors")
  }
  
  mean_x = mean(x)
  sd_x = sd(x)

  out_df =
    tibble(
      mean = mean_x,
      sd = sd_x
    )
  return (out_df)
}

mean_and_sd (x_vec)
```

# Check stuff using a simulation
```{r}
sim_df = 
  tibble(
  x = rnorm(30, mean = 10, sd = 5)
)

# traditional method:
sim_df |> 
  summarize(
    mean = mean(x),
    sd = sd(x)
  )
```

Simulation function to check sample mean and sd:
```{r}
sim_mean_sd = function (sample_size, true_mean, true_sd) {
  
  sim_df = 
  tibble(
  x = rnorm(sample_size, true_mean, true_sd)
)

out_df =
  sim_df |> 
  summarize(
    mean = mean(x),
    sd = sd(x)
  )
  
return(out_df)
}

sim_mean_sd (sample_size = 30, true_mean = 4, true_sd =12)
sim_mean_sd (30, 4, 12)

# "sim_mean_sd (300)" # the function doesn't work until we do the following:

sim_mean_sd = function (sample_size = 30, true_mean =10, true_sd = 5) {
  
  sim_df = 
  tibble(
  x = rnorm(sample_size, true_mean, true_sd)
)

out_df =
  sim_df |> 
  summarize(
    mean = mean(x),
    sd = sd(x)
  )
  
return(out_df)
}

sim_mean_sd (sample_size = 30, true_mean = 4, true_sd =12)

sim_mean_sd (300)
```


```{r}
fellowship_ring = readxl::read_excel("./data/LotR_Words.xlsx", range = "B3:D6") |>
  mutate(movie = "fellowship_ring")

two_towers = readxl::read_excel("./data/LotR_Words.xlsx", range = "F3:H6") |>
  mutate(movie = "two_towers")

return_king = readxl::read_excel("./data/LotR_Words.xlsx", range = "J3:L6") |>
  mutate(movie = "return_king")

lotr_tidy = bind_rows(fellowship_ring, two_towers, return_king) |>
  janitor::clean_names() |>
  pivot_longer(
    female:male,
    names_to = "sex",
    values_to = "words") |> 
  mutate(race = str_to_lower(race)) |> 
  select(movie, everything()) 
```

# Try to write a function to import the data above
```{r}
lotr_import = function (cell_range, movie_title) {
  
movie_df = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = cell_range) |>
  mutate (movie = movie_title) |> 
  janitor::clean_names() |> 
  pivot_longer(
      female:male,
      names_to = "sex",
      values_to = "words") |>
    mutate(
      race = str_to_lower(race),
      movie = movie_title)
  
return (movie_df)
}

lotr_df = 
  bind_rows(
lotr_import ("B3:D6", "fellowship_ring"),
lotr_import ("F3:H6", "two_towers"),
lotr_import ("J3:L6", "return_king"),
)
```


```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

nsduh_html = read_html(nsduh_url)

data_marj = 
  nsduh_html |> 
  html_table() |> 
  nth(1) |> # nth(4) and nth(5) to take No. 4 and No. 5 tables
  slice(-1) |> 
  select(-contains("P Value")) |>
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") |>
  separate(age_year, into = c("age", "year"), sep = "\\(") |>
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)) |>
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
```



# Let’s write a quick function to scrape review information for other tables on this page. We’ll pass in the HTML data as an argument so we don’t scrape it each time, along with a number and name for the table we want to process.

```{r}
nsduh_table_format = function (table_num, table_name) {
  out_table =
   nsduh_html |> 
    html_table() |> 
    nth(table_num) |> 
    slice (-1) |> 
    mutate (name = table_name)
  
  return (out_table)
}

nsduh_table_format (1, "marj")
```

# or we can do some cleaning as follows:
```{r}
nsduh_table = function(html, table_num, table_name) {
  
  table = 
    html |> 
    html_table() |> 
    nth(table_num) |>
    slice(-1) |> 
    select(-contains("P Value")) |>
    pivot_longer(
      -State,
      names_to = "age_year", 
      values_to = "percent") |>
    separate(age_year, into = c("age", "year"), sep = "\\(") |>
    mutate(
      year = str_replace(year, "\\)", ""),
      percent = str_replace(percent, "[a-c]$", ""),
      percent = as.numeric(percent),
      name = table_name) |>
    filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
  
  return (table)
  
}
```

```{r}
nsduh_results = 
  bind_rows(
    nsduh_table(nsduh_html, 1, "marj_one_year"),
    nsduh_table(nsduh_html, 4, "cocaine_one_year"),
    nsduh_table(nsduh_html, 5, "heroin_one_year")
  )
```

```{r}
x_vec = rnorm(25, 0, 1)

my_summary = function(x, summ_func) {
  summ_func(x)
}

my_summary(x_vec, sd)

my_summary(x_vec, IQR)

my_summary(x_vec, var)
```


```{r}
f = function(x) {
  z = x + y
  z
}

x = 1
y = 2

f(x = y)
```

